---
title: "Airline Pricing Analysis"
author: "Vaibhav Srivastav"
date: "16 February 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Airline Pricing Analysis

This is a mini project on the analysis of the difference between the pricing of premium economy and economy class airline tickets.

## 1. Set the working directory.

```{r wd}
setwd("~/Desktop/Data Analytics Internship/Airline")
```

## 2. Read the dataset using the read.csv() function.

```{r rd}
air<-read.csv(file="SixAirlinesDataV2.csv")
```

## 3. Use the View() function to view the dataframe created in the previous step.

```{r view}
View(air)
```

## 4. Describing the Dataset

```{r desc}
library(psych)
describe(air)
```

# Summary

```{r summ}
summary(air)
```

# Structure of Dataset

```{r struc}
str(air)  
```

## Plots to visualize the distribution of each variable independently

# Flight Duration

```{r plot1}
boxplot(air$FlightDuration, horizontal=TRUE,
        main="Flight duration of all the airlines")
```

# Economic Seats

```{r plot2}
boxplot(air$SeatsEconomy, horizontal=TRUE,
        main="Economic seats of all airlines")
```

# Premium seats

```{r plot3}
boxplot(air$SeatsPremium, horizontal=TRUE,
        main="Flight durations of all the airlines")
```

# Economy Price

```{r plot4}
boxplot(air$PriceEconomy, horizontal=TRUE,
        main="Economic prices of all airlines")
```

# Premium Price

```{r plot5}
boxplot(air$PricePremium, horizontal=TRUE,
        main="Premium prices of all airlines")
```

## Boxplot for relative price for each pitch difference

```{r relative}
boxplot(air$PriceRelative ~ air$PitchDifference, horizontal=TRUE,
        xlab = "Price Relative", ylab = "Pitch Difference",
        main = "Relative price changes according to pitch difference",
        col=c("red","blue","green","yellow","brown")
          )
```

## 

```{r relative2}
boxplot(air$PriceRelative ~ air$WidthDifference, horizontal=TRUE,
        xlab = "Price Relative", ylab = "Width Difference",
        main = "Relative price changes according to width difference",
        col=c("red","blue","green","yellow","brown")
          )
```

## Scatter Plots to understand how are the variables correlated pair-wise

# Economic pricing variables
```{r scatplot}
library(car)
 scatterplotMatrix(formula = ~ FlightDuration + SeatsEconomy + PitchEconomy + WidthEconomy + PriceEconomy + PriceRelative, data = air, diagonal="histogram")
```

# Premium pricing dependent variables

```{r scatplot2}
scatterplotMatrix(formula = ~ FlightDuration + SeatsPremium + PitchPremium + WidthPremium + PricePremium + PriceRelative, data = air, diagonal="histogram")
```

## Corrgram
# Create a Variance-Covariance Matrix


```{r corg}
library(corrgram)
corrgram(air, order=FALSE, 
         lower.panel=panel.shade,
         upper.panel=panel.pie, 
         diag.panel=panel.minmax,
         text.panel=panel.txt,
         main="Corrgram of airlines intercorrelations")
```

## Correlation

```{r cor}
cor.test(air$PriceRelative, air$PitchDifference)

```
As p<0.05, these 2 variables are strongly corelated.

```{r cor2}
cor.test(air$PriceRelative, air$WidthDifference)
```
As p<0.05, these 2 variables are strongly corelated.

## Regression Model

Pitchpremium,widthpremium,pitchdifference,pricerelative,widtheconomy,flightduration,seatseconomy has a strong positive correlation with one another. Also, There is a strong positive correlation between pitchpremium and widthpremium , widthpremium and width difference, etc.

# Premium Pricing 

Dependent variable: PricePremium 
Independent variable: Flightduration,PitchPremium and WidthPremium 
Independent variables are affecting/influencing the Pricing of Premium class tickets.

Following linear regression model considered,
*PricePremium = x0 + x1FlightDuration + x2WidthPremium + x3PitchPremium*

```{r prem}
prem <- lm(air$PricePremium~air$FlightDuration+air$WidthPremium+air$PitchPremium)
summary(prem)
```
There is a significant effect of width and pitch on the pricing of the premium class tickets. Now, assuming the same for economic class.

# Economic Pricing 

Dependent variable: PriceEconomy
Independent variable: Flightduration,PitchEconomy and WidthEconomy
Independent variables are affecting/influencing the Pricing of Economy class tickets.

```{r eco}
eco <- lm(air$PriceEconomy~air$FlightDuration+air$WidthEconomy+air$PitchEconomy)
summary(eco)
```
Hence, Flight duration, width of the seat and the pitch of the seat are the factors that influence the pricing of airline tickets.
